\chapter{Introduction}

The Standard Model (SM) of particle physics is our best descriptor for the behaviour 
of elementary particles and the interactions between them.
Even if it motivated major successes in particle physics, like the finding of the Higgs boson or the top quark, many phenomena stay unexplained. 
Physics beyond the Standard Model (BSM) is needed to make breakthroughs in 
dark matter, neutrino oscillations or the matter-antimattter disbalance. 

As the heaviest fermion with a mass of about \SI{172.5}{\GeV}, the top quark couples 
the strongest with the Higgs field and may act as a potential probe for electroweak 
symmetry breaking. Therefore there lies potential in uncovering BSM physics by further
studying it. 
The challenge lies in its wide decay width and short life time of \SI{5e-25}{\s}, 
which makes hadronisation impossible and almost exclusively leads to a $W$-boson and
bottom-quark as decay products.   

At the ATLAS experiment of the Large Hadron Collider (LHC) top-quarks predominantly 
come in pairs of $\ttbar$ via the strong interaction.  
Although, top quarks can be produced singly via the electroweak interaction, with 
the single top production in the s- and t-channel, aswell as in $tW$ associated production in leading order (LO). Each of these modes produces the top-quark in 
a $tWb$ vertex, 


In each single top-quark production mode a top quark is produced with a $tWb$ vertex, where the $V_{\rm{tb}}$ element 
of the Cabbibo-Kobayashi-Maskawa (CKM) matrix is involved. 
The s-channel has the smallest cross section of \SI{10.32}{\piko\barn} off all mentionend processes and is dominated by background processes.
Unlike the other single modes, the s-channel
has not yet been measured precisely with a significance over $5 \sigma$.
The latest ATLAS measurement achieved an observed (expected) significance of 3.3(3.9) standard deviations, leaving missing piece of picture (wieso kann Niklas so gut schreiben maaaan ey).  
Even with its small cross section, the major hindrances in the measurement are the systematic uncertainties. To get a better measurement this work will use a profile likelihood fit, with a DNN output as discriminant, on the full Run 2 dataset collected with a lumi of \SI{140}{\per\femto\barn}

Do not forget structur of thesis
